---
title: "{{STUDY_NAME}} — 00b Pipeline Status"
subtitle: "What is built, what is approved, what comes next"
date: today
format:
  html:
    toc: true
    code-fold: true
    code-summary: "Show code"
    embed-resources: true
    df-print: kable
execute:
  warning: false
  message: false
---

## Overview

The build record. `_pipeline.yml` holds one entry per stage with its status
(planned, generated, rendered, approved, failed), and this notebook draws it:
a diagram of the stages and their dependencies, coloured by status, and the
table underneath. Re-render it whenever you want to see where the build
stands; it reads nothing but the manifest and the coverage files the modality
stages leave in `output/validation/`.

Two rules are enforced by the manifest and are worth knowing before reading
the picture. A stage is generated only after every stage it depends on is
approved, and a stage is approved only after its render has been looked at.
So a blue node has been seen by a person; a grey node downstream of it has
not been started; a red node stopped and its note says why.

```{r}
#| label: setup
suppressPackageStartupMessages({ library(fearlabr); library(dplyr) })
config <- read_config()
stem <- fearlabr_file_stem(config)
builder_stage_gate("00b_pipeline_status")
manifest <- read_pipeline_manifest()
validation_dir <- resolve_path(config$paths$validation %||% "output/validation", config)
figures_dir <- resolve_path(config$paths$figures %||% "output/figures", config)
dir.create(figures_dir, recursive = TRUE, showWarnings = FALSE)
```

## The pipeline

```{r}
#| label: dag
#| fig-width: 11
#| fig-height: 5
log_section("Pipeline DAG")
cat("• Study: ", manifest$study, "   modalities: ", paste(unlist(manifest$modalities), collapse = ", "),
    "   timepoints: ", paste(unlist(manifest$timepoints), collapse = ", "), "\n", sep = "")
plot_pipeline_dag(manifest)

# The same graph as Mermaid text, for the README and for the Shiny front end.
mmd_path <- file.path(dirname(validation_dir), "pipeline_dag.mmd")
writeLines(pipeline_dag_mermaid(manifest), mmd_path)
cat("✓ Mermaid source written: ", mmd_path, "\n", sep = "")
```

## Status table

```{r}
#| label: status
status <- pipeline_status_table(manifest)
n_by <- table(factor(status$status, levels = stage_status_levels()))
cat("• ", paste(names(n_by), n_by, sep = ": ", collapse = "   "), "\n", sep = "")
nxt <- next_stage(manifest)
cat(if (is.null(nxt)) "✓ Every stage is approved.\n" else paste0("• Next stage to build: ", nxt, "\n"))
status
```

## Coverage across modalities

Each modality stage writes `output/validation/<stem>_coverage_<modality>.csv`
(participants expected at each timepoint, participants observed). When at
least one exists they are drawn together here, so a wave where EEG has 40 of
60 while REDCap has 58 of 60 is visible before anyone models it.

```{r}
#| label: coverage
#| fig-width: 10
#| fig-height: 3.5
cov_files <- list.files(validation_dir, pattern = paste0("^", stem, "_coverage_.*\\.csv$"), full.names = TRUE)
if (length(cov_files)) {
  dash <- modality_coverage_dashboard(lapply(cov_files, readr::read_csv, show_col_types = FALSE))
  cat("✓ ", length(cov_files), " coverage file(s): ", paste(basename(cov_files), collapse = ", "), "\n", sep = "")
  print(plot_coverage_dashboard(dash))
  dash
} else {
  cat("⏭ No coverage files yet; they appear once a modality's QC stage has rendered\n")
}
```

## What happens next

Approve this stage in the R console with `builder_approve("00b_pipeline_status")`
once the diagram matches the plan you agreed to. The next stage is whichever
the status table names; it is generated only after this one is approved.

```{r}
#| label: record
builder_record_render("00b_pipeline_status")
```
