---
title: "{{STUDY_NAME}} — 01e Ingest EEG features"
subtitle: "Preprocessed feature export → canonical long table in the study's ID and timepoint vocabulary"
date: today
format:
  html:
    toc: true
    code-fold: true
    code-summary: "Show code"
    embed-resources: true
    df-print: kable
execute:
  warning: true
  message: false
---

## Overview

The EEG that reaches this pipeline has already been preprocessed. Filtering,
epoching, artifact rejection and averaging were done in the EEG tool
(MNE, EEGLAB, ERPLAB or BrainVision Analyzer), and what it exported is a long
table: one row per participant, session, channel, condition and measure.
This stage does not touch the signal. It does three things: renames the
export's columns to the pipeline's names through the crosswalk in
`eeg.columns`, brings the participant IDs into the study's format
(`sub-0012` becomes `0012`), and maps the export's session labels to the
declared timepoints through `eeg.session_map`.

Run it after 00 and before 02e. Nothing here is scored or gated on quality;
02e does that on the file this stage writes.

```{r}
#| label: setup
suppressPackageStartupMessages({ library(fearlabr); library(dplyr) })
config <- read_config()
stem <- fearlabr_file_stem(config)
builder_stage_gate("01e_ingest_eeg")
raw_dir <- resolve_path(config$paths$raw_data, config)
validation_dir <- resolve_path(config$paths$validation %||% "output/validation", config)
figures_dir <- resolve_path(config$paths$figures %||% "output/figures", config)
for (d in c(validation_dir, figures_dir)) dir.create(d, recursive = TRUE, showWarnings = FALSE)
```

## Which files

```{r}
#| label: files
log_section("EEG export files")
files <- builder_glob_files(config, config$eeg$file_glob)
if (!length(files)) stop("No EEG export matches eeg.file_glob '", config$eeg$file_glob,
                         "' under ", raw_dir, ". Drop the export there and re-render.")
cat("• ", length(files), " file(s): ", paste(basename(files), collapse = ", "), "\n", sep = "")
cat("• Crosswalk: ", paste(names(config$eeg$columns), unlist(config$eeg$columns),
                           sep = " <- ", collapse = ", "), "\n", sep = "")
```

## Ingest

The ID print below is the line to read. `distinct` is the number of people in
the export; `format-fail` is how many IDs do not match the study's pattern
after the prefix strip and padding. A nonzero format-fail is a crosswalk or
transform problem, not a data problem, and it will under-join later.

```{r}
#| label: ingest
log_section("Ingest")
eeg_long <- ingest_eeg_features(config, files = files, write = TRUE)
cat("✓ ", nrow(eeg_long), " rows; ", dplyr::n_distinct(eeg_long$study_id), " participants; ",
    dplyr::n_distinct(eeg_long$channel), " channels; ",
    dplyr::n_distinct(eeg_long$measure), " measure(s)\n", sep = "")
cat("✓ Wrote ", stem, "_eeg_features_latest.rds to ", raw_dir, "\n", sep = "")
```

## Session to timepoint

Every export session label must map to a declared timepoint. A session that
maps to nothing is kept with `timepoint = NA` so it shows up here and in 02e,
rather than vanishing.

```{r}
#| label: sessions
eeg_long |>
  count(eeg_session, timepoint, name = "n_rows") |>
  mutate(n_participants = purrr::map2_int(eeg_session, timepoint, ~ n_distinct(
    eeg_long$study_id[eeg_long$eeg_session %in% .x & (is.na(.y) | eeg_long$timepoint %in% .y)])))
```

## Quick look: values by measure and condition

Before any QC rule runs, the raw distributions. Look for a measure whose
values sit on the wrong scale (microvolts exported as volts show up as a
band around zero) or a condition present in one session and absent from
another.

```{r}
#| label: quicklook
#| fig-width: 10
#| fig-height: 4
if (requireNamespace("ggplot2", quietly = TRUE)) {
  p <- ggplot2::ggplot(eeg_long, ggplot2::aes(x = timepoint, y = value, fill = condition)) +
    ggplot2::geom_boxplot(outlier.size = 0.6) +
    ggplot2::facet_wrap(~ measure, scales = "free_y") +
    ggplot2::labs(x = NULL, y = "Exported value") +
    ggplot2::theme_minimal(base_size = 11)
  ggplot2::ggsave(file.path(figures_dir, paste0(stem, "_eeg_raw_values.png")), p, width = 10, height = 4, dpi = 120)
  print(p)
}
eeg_long |>
  summarise(n = n(), min = min(value, na.rm = TRUE), median = median(value, na.rm = TRUE),
            max = max(value, na.rm = TRUE), n_missing = sum(is.na(value)), .by = c(measure, condition))
```

## Coverage against the roster

Expected participants come from the REDCap participant-level file when 03 has
run; otherwise from the export itself, which can only show gaps within the
people who have EEG at all. The coverage file feeds 00b and 07m.

```{r}
#| label: coverage
#| fig-width: 8
#| fig-height: 6
roster_path <- file.path(resolve_path(config$paths$derived, config),
                         paste0(stem, "_redcap_participant_level_latest.rds"))
expected <- if (file.exists(roster_path)) {
  r <- readRDS(roster_path); as.character(r[[config$redcap$id_column]])
} else {
  cat("⚠️ No REDCap participant-level file yet; coverage is relative to the EEG export only\n")
  unique(eeg_long$study_id)
}
cov <- timepoint_coverage(eeg_long, "study_id", "timepoint", config, modality = "eeg",
                          expected_ids = expected)
cov_summary <- timepoint_coverage_summary(cov)
readr::write_csv(cov_summary, file.path(validation_dir, paste0(stem, "_coverage_eeg.csv")))
cat("✓ Coverage written: ", stem, "_coverage_eeg.csv\n", sep = "")
if (requireNamespace("ggplot2", quietly = TRUE)) print(plot_timepoint_completeness(cov))
cov_summary
```

## What happens next

02e reads `<stem>_eeg_features_latest.rds`, applies the QC rules in
`eeg.qc`, and reduces the long table to the features named in
`eeg.features`. Approve this stage once the ID line shows zero format
failures, every session maps to a timepoint you recognise, and the value
scales look like the units the export claims.

```{r}
#| label: record
builder_record_render("01e_ingest_eeg")
```
