---
title: "{{STUDY_NAME}} — 01s Ingest sensor streams"
subtitle: "Vendor exports → canonical participant-days, one table per stream"
date: today
format:
  html:
    toc: true
    code-fold: true
    code-summary: "Show code"
    embed-resources: true
    df-print: kable
execute:
  warning: true
  message: false
---

## Overview

Each stream declared under `sensors.streams` (a wrist actigraph's daily
summary, a phone's accelerometer epochs, a sleep export) is read from the
files its glob matches, its vendor columns are renamed to the metric names
the study uses, and epoch rows are collapsed to one row per participant-day
with the aggregate the config declares. Dates are local calendar dates in
the configured time zone; an epoch timestamp is never turned into a date
through `as.Date()` on a POSIXct, which ignores the time zone and moved a
fifth of one study's evening events onto the next day.

Nothing is judged here. Which days count as valid is 02s's job, by the rule
the config states.

```{r}
#| label: setup
suppressPackageStartupMessages({ library(fearlabr); library(dplyr) })
config <- read_config()
stem <- fearlabr_file_stem(config)
builder_stage_gate("01s_ingest_sensor")
raw_dir <- resolve_path(config$paths$raw_data, config)
figures_dir <- resolve_path(config$paths$figures %||% "output/figures", config)
dir.create(figures_dir, recursive = TRUE, showWarnings = FALSE)
streams <- names(config$sensors$streams)
cat("• Streams: ", paste(streams, collapse = ", "), "   tz: ", config$sensors$tz %||% "UTC", "\n", sep = "")
```

## Ingest, stream by stream

Read each block's ID line. `dup` is expected to be large here (one row per
day per person); `format-fail` and `missing` are the numbers that matter.

```{r}
#| label: ingest
day_tables <- list()
for (sk in streams) {
  log_section(paste0("Stream: ", sk, " (", config$sensors$streams[[sk]]$label %||% sk, ")"))
  files <- builder_glob_files(config, config$sensors$streams[[sk]]$file_glob)
  if (!length(files)) {
    cat("⏭ No file matches '", config$sensors$streams[[sk]]$file_glob, "'; stream skipped\n", sep = "")
    next
  }
  cat("• ", length(files), " file(s): ", paste(basename(files), collapse = ", "), "\n", sep = "")
  day <- ingest_sensor_stream(config, sk, files = files, write = TRUE)
  cat("✓ ", nrow(day), " participant-days; ", dplyr::n_distinct(day$study_id), " participants; ",
      format(min(day$date)), " to ", format(max(day$date)), "\n", sep = "")
  cat("✓ Wrote ", stem, "_sensor_", sk, "_day_latest.rds\n", sep = "")
  day_tables[[sk]] <- day
}
if (!length(day_tables)) stop("No sensor stream had files; drop exports under ", raw_dir, " and re-render.")
```

## Quick look: observed days per participant

Before any validity rule, simply which days exist. A participant whose block
starts weeks after everyone else's, or a stream that ends on the same date
for everyone, is a device or export problem to raise now.

```{r}
#| label: quicklook
#| fig-width: 10
#| fig-height: 6
for (sk in names(day_tables)) {
  d <- day_tables[[sk]]
  cat("• ", sk, ": median ", median(table(d$study_id)), " days per participant\n", sep = "")
  if (requireNamespace("ggplot2", quietly = TRUE)) {
    p <- plot_sensor_coverage(d |> mutate(valid_day = TRUE)) +
      ggplot2::labs(title = paste0(sk, ": observed days")) + ggplot2::theme(legend.position = "none")
    ggplot2::ggsave(file.path(figures_dir, paste0(stem, "_sensor_", sk, "_observed.png")), p,
                    width = 10, height = 6, dpi = 120)
    print(p)
  }
}
bind_rows(lapply(day_tables, function(d) summarise(d, n_participants = n_distinct(study_id),
                                                     n_days = n(), first = min(date), last = max(date),
                                                     .by = sensor_stream)))
```

## What happens next

02s applies each stream's `valid_day_rule`, reports coverage per participant,
and attaches the declared timepoints using each participant's anchor date
from REDCap. Approve this stage once every stream you expected was read,
the ID lines show zero format failures, and the date ranges match the
study's calendar.

```{r}
#| label: record
builder_record_render("01s_ingest_sensor")
```
