---
title: "{{STUDY_NAME}} — 02e EEG QC and features"
subtitle: "Trial floor, amplitude range, required channels → feature table per participant × timepoint"
date: today
format:
  html:
    toc: true
    code-fold: true
    code-summary: "Show code"
    embed-resources: true
    df-print: kable
execute:
  warning: true
  message: false
---

## Overview

Three QC rules, each read from `eeg.qc` and printed here so the render
documents what was applied: a minimum trial count per participant and
session, an amplitude range outside which a value is implausible, and the
channels that must be present. A participant-timepoint failing any rule is
flagged, not dropped; the flag travels with the feature table so an analysis
can exclude on it explicitly.

Then the reduction: each feature in `eeg.features` is the mean of its measure
over its channels for its condition. A feature that matches no rows stops
this stage and names itself. The alternative, a column of NA that looks like
missing data, cost a study a week once.

```{r}
#| label: setup
suppressPackageStartupMessages({ library(fearlabr); library(dplyr) })
config <- read_config()
stem <- fearlabr_file_stem(config)
builder_stage_gate("02e_qc_eeg")
raw_dir <- resolve_path(config$paths$raw_data, config)
derived_dir <- resolve_path(config$paths$derived, config)
validation_dir <- resolve_path(config$paths$validation %||% "output/validation", config)
figures_dir <- resolve_path(config$paths$figures %||% "output/figures", config)
for (d in c(derived_dir, validation_dir, figures_dir)) dir.create(d, recursive = TRUE, showWarnings = FALSE)

eeg_long <- readRDS(file.path(raw_dir, paste0(stem, "_eeg_features_latest.rds")))
id_audit_line(eeg_long$study_id, "EEG long (from 01e)", pattern = config$eeg$id_pattern)
```

## The rules, as configured

```{r}
#| label: rules
qc_cfg <- config$eeg$qc
cat("• min_trials:         ", qc_cfg$min_trials %||% "not set", "\n", sep = "")
cat("• amplitude_range_uv: ", paste(qc_cfg$amplitude_range_uv %||% c("-Inf", "Inf"), collapse = " to "), "\n", sep = "")
cat("• required_channels:  ", paste(qc_cfg$required_channels %||% "none", collapse = ", "), "\n", sep = "")
```

## QC verdicts

```{r}
#| label: qc
#| fig-width: 8
#| fig-height: 6
log_section("QC")
qc <- eeg_qc_summary(eeg_long, config)
n_flag <- sum(!qc$pass)
cat(if (n_flag == 0) "✓ Every participant-timepoint passes\n" else
    paste0("⚠️ ", n_flag, " of ", nrow(qc), " participant-timepoint(s) flagged\n"))
readr::write_csv(qc, file.path(validation_dir, paste0(stem, "_eeg_qc.csv")))
if (requireNamespace("ggplot2", quietly = TRUE)) {
  p <- plot_eeg_qc(qc)
  ggplot2::ggsave(file.path(figures_dir, paste0(stem, "_eeg_qc.png")), p, width = 8, height = 6, dpi = 120)
  print(p)
}
qc |> filter(!pass)
```

## Feature table

```{r}
#| label: features
#| fig-width: 10
#| fig-height: 4
log_section("Features")
features <- eeg_feature_table(eeg_long, config) |>
  left_join(qc |> select(study_id, timepoint, eeg_qc_pass = pass), by = c("study_id", "timepoint"))
assert_expected_n(features, dplyr::n_distinct(eeg_long$study_id), id_col = "study_id",
                  label = "feature reduction")
write_versioned_rds(features, paste0(stem, "_eeg_features_wide"), derived_dir)
write_latest_rds(features, paste0(stem, "_eeg_features_wide"), derived_dir)
cat("✓ ", nrow(features), " participant-timepoint rows; features: ",
    paste(vapply(config$eeg$features, function(f) f$name, character(1)), collapse = ", "), "\n", sep = "")
if (requireNamespace("ggplot2", quietly = TRUE)) {
  p <- plot_eeg_features(features |> select(-eeg_qc_pass))
  ggplot2::ggsave(file.path(figures_dir, paste0(stem, "_eeg_features.png")), p, width = 10, height = 4, dpi = 120)
  print(p)
}
features |>
  tidyr::pivot_longer(-c(study_id, timepoint, eeg_qc_pass), names_to = "feature") |>
  summarise(n = sum(!is.na(value)), mean = mean(value, na.rm = TRUE), sd = sd(value, na.rm = TRUE),
            .by = c(feature, timepoint))
```

## What happens next

`<stem>_eeg_features_wide_latest.rds` is what 07m links to REDCap and the
other modalities, on `study_id` and `timepoint`. Approve this stage when the
flagged list is one you can account for (a known bad session, a participant
who left early) and the feature distributions are on the scale the literature
reports for that component.

```{r}
#| label: record
builder_record_render("02e_qc_eeg")
```
