---
title: "{{STUDY_NAME}} — 02s Sensor coverage"
subtitle: "Valid days by the declared rule, coverage per participant, days assigned to timepoints"
date: today
format:
  html:
    toc: true
    code-fold: true
    code-summary: "Show code"
    embed-resources: true
    df-print: kable
execute:
  warning: true
  message: false
---

## Overview

A day of actigraphy with the watch on the nightstand is not a day of data.
Each stream's `valid_day_rule` in the config says what counts (for wrist
actigraphy, commonly at least 600 wear minutes), and this stage prints the
rule verbatim, applies it, and shows the result as a participant-by-day tile
so a device that stopped recording is visible as a run of orange.

Days are then assigned to the declared timepoints using each participant's
anchor date from REDCap (`timepoints.anchor_date_column`), so sensor coverage
can be reported in the same wave vocabulary as everything else.

```{r}
#| label: setup
suppressPackageStartupMessages({ library(fearlabr); library(dplyr) })
config <- read_config()
stem <- fearlabr_file_stem(config)
builder_stage_gate("02s_qc_sensor")
raw_dir <- resolve_path(config$paths$raw_data, config)
derived_dir <- resolve_path(config$paths$derived, config)
validation_dir <- resolve_path(config$paths$validation %||% "output/validation", config)
figures_dir <- resolve_path(config$paths$figures %||% "output/figures", config)
for (d in c(derived_dir, validation_dir, figures_dir)) dir.create(d, recursive = TRUE, showWarnings = FALSE)
streams <- names(config$sensors$streams)
```

## Anchor dates

```{r}
#| label: anchors
roster_path <- file.path(derived_dir, paste0(stem, "_redcap_participant_level_latest.rds"))
anchor_col <- config$timepoints$anchor_date_column
anchors <- if (file.exists(roster_path) && !is.null(anchor_col)) {
  r <- readRDS(roster_path)
  if (!anchor_col %in% names(r)) stop("timepoints.anchor_date_column '", anchor_col,
                                      "' is not in the participant-level file.")
  tibble(study_id = as.character(r[[config$redcap$id_column]]), anchor_date = as.Date(r[[anchor_col]]))
} else {
  cat("⚠️ No participant-level file or no anchor_date_column; timepoints cannot be assigned yet\n")
  NULL
}
if (!is.null(anchors)) {
  id_audit_line(anchors$study_id, "REDCap anchors")
  cat("• ", sum(is.na(anchors$anchor_date)), " participant(s) with a missing anchor date\n", sep = "")
}
```

## Streams

```{r}
#| label: streams
#| fig-width: 10
#| fig-height: 6
coverage_all <- list(); summaries <- list()
for (sk in streams) {
  f <- file.path(raw_dir, paste0(stem, "_sensor_", sk, "_day_latest.rds"))
  if (!file.exists(f)) { cat("⏭ ", sk, ": no day table from 01s\n", sep = ""); next }
  log_section(paste0("Stream: ", sk))
  rule <- config$sensors$streams[[sk]]$valid_day_rule
  cat("• valid_day_rule: ", rule, "\n", sep = "")
  day <- readRDS(f) |> sensor_valid_days(rule)
  cat("✓ ", sum(day$valid_day), " of ", nrow(day), " participant-days valid (",
      round(100 * mean(day$valid_day), 1), "%)\n", sep = "")
  if (requireNamespace("ggplot2", quietly = TRUE)) {
    p <- plot_sensor_coverage(day) + ggplot2::labs(title = paste0(sk, ": ", rule))
    ggplot2::ggsave(file.path(figures_dir, paste0(stem, "_sensor_", sk, "_valid.png")), p,
                    width = 10, height = 6, dpi = 120)
    print(p)
  }
  cov <- sensor_coverage(day)
  readr::write_csv(cov, file.path(validation_dir, paste0(stem, "_sensor_", sk, "_coverage.csv")))
  summaries[[sk]] <- cov |> mutate(sensor_stream = sk, .before = 1)

  if (!is.null(anchors)) {
    day <- link_sensor_to_timepoints(day, anchors, config)
    cat("• ", sum(!is.na(day$timepoint)), " of ", nrow(day), " days fall inside a declared window\n", sep = "")
    coverage_all[[sk]] <- timepoint_coverage(day |> filter(valid_day), "study_id", "timepoint", config,
                                             modality = "sensors", expected_ids = anchors$study_id) |>
      mutate(sensor_stream = sk)
  }
  write_versioned_rds(day, paste0(stem, "_sensor_", sk, "_day_valid"), derived_dir)
  write_latest_rds(day, paste0(stem, "_sensor_", sk, "_day_valid"), derived_dir)
  cat("✓ Wrote ", stem, "_sensor_", sk, "_day_valid_latest.rds\n", sep = "")
}
bind_rows(summaries) |>
  summarise(n_participants = n(), median_valid_days = median(n_valid_days),
            min_valid_days = min(n_valid_days), pct_under_7_valid = round(100 * mean(n_valid_days < 7), 1),
            .by = sensor_stream)
```

## Coverage by timepoint

Participants counted as observed at a timepoint have at least one valid day
inside that timepoint's window. The file written here feeds 00b and 07m.

```{r}
#| label: coverage
#| fig-width: 8
#| fig-height: 6
if (length(coverage_all)) {
  cov <- bind_rows(coverage_all)
  # One modality row per timepoint: observed if any stream has a valid day there.
  cov_any <- cov |> summarise(observed = any(observed), n_obs = sum(n_obs),
                              .by = c(study_id, timepoint, modality))
  cov_summary <- timepoint_coverage_summary(cov_any)
  readr::write_csv(cov_summary, file.path(validation_dir, paste0(stem, "_coverage_sensors.csv")))
  cat("✓ Coverage written: ", stem, "_coverage_sensors.csv\n", sep = "")
  if (requireNamespace("ggplot2", quietly = TRUE)) print(plot_timepoint_completeness(cov_any))
  cov_summary
} else {
  cat("⏭ Timepoint coverage skipped: anchors unavailable (run 03 first, or set timepoints.anchor_date_column)\n")
}
```

## What happens next

07m joins the valid-day tables to REDCap and the other modalities. Approve
this stage when the rule printed above is the one the protocol states, the
tile shows the gaps you already know about and no others, and the
per-participant table has no one with an implausible valid-day count.

```{r}
#| label: record
builder_record_render("02s_qc_sensor")
```
