---
title: "{{STUDY_NAME}} — 07m Link modalities"
subtitle: "Every modality on one participant × timepoint index, with the ID audit across all of them"
date: today
format:
  html:
    toc: true
    code-fold: true
    code-summary: "Show code"
    embed-resources: true
    df-print: kable
execute:
  warning: true
  message: false
---

## Overview

The first stage where the modalities meet. It builds one index, one row per
participant per declared timepoint, with a column per modality saying whether
that modality was observed there, then joins the modality tables onto it.
Before the join it runs the ID audit across every source at once, because a
format that drifted in one export (padding, a prefix, a case change) is the
mechanism behind a join that quietly reaches a third of the sample.

`study_id` is the key everywhere and is character. The join is `study_id`
and `timepoint`; a modality without timepoints (a participant-level table) joins on
`study_id` alone and is repeated across timepoints, which is the intended
meaning of a baseline covariate.

```{r}
#| label: setup
suppressPackageStartupMessages({ library(fearlabr); library(dplyr) })
config <- read_config()
stem <- fearlabr_file_stem(config)
builder_stage_gate("07m_link_modalities")
derived_dir <- resolve_path(config$paths$derived, config)
validation_dir <- resolve_path(config$paths$validation %||% "output/validation", config)
figures_dir <- resolve_path(config$paths$figures %||% "output/figures", config)
for (d in c(validation_dir, figures_dir)) dir.create(d, recursive = TRUE, showWarnings = FALSE)
id_col <- config$redcap$id_column
modalities <- config_modalities(config)
sched <- timepoint_schedule(config)
```

## Gather the sources

```{r}
#| label: gather
log_section("Sources")
latest <- function(name) {
  f <- file.path(derived_dir, paste0(stem, "_", name, "_latest.rds"))
  if (file.exists(f)) readRDS(f) else NULL
}
roster <- latest("redcap_participant_level")
if (is.null(roster)) stop("03_clean_redcap has not written the participant-level file; approve 03 first.")
roster$study_id <- as.character(roster[[id_col]])
sources <- list(REDCap = roster$study_id)
obs <- list()

# REDCap assessments carry the timepoint through the event -> timepoint map.
assess <- latest("redcap_assessment_long")
if (!is.null(assess) && "redcap_event_name" %in% names(assess)) {
  ev_map <- setNames(sched$timepoint, sched$redcap_event)
  assess$study_id <- as.character(assess[[id_col]])
  assess$timepoint <- unname(ev_map[assess$redcap_event_name])
  obs$redcap <- assess |> filter(!is.na(timepoint)) |> distinct(study_id, timepoint)
}
ema <- latest("ema_week_linked") %||% latest("ema_event_analysis")
if ("ema" %in% modalities && !is.null(ema) && "study_id" %in% names(ema)) {
  sources$EMA <- ema$study_id
  if ("timepoint" %in% names(ema)) obs$ema <- ema |> distinct(study_id, timepoint)
}
eeg <- latest("eeg_features_wide")
if ("eeg" %in% modalities && !is.null(eeg)) {
  sources$EEG <- eeg$study_id
  obs$eeg <- eeg |> distinct(study_id, timepoint)
}
if ("sensors" %in% modalities) {
  for (sk in names(config$sensors$streams)) {
    d <- latest(paste0("sensor_", sk, "_day_valid"))
    if (is.null(d)) next
    sources[[paste0("sensor:", sk)]] <- d$study_id
    if ("timepoint" %in% names(d)) {
      obs$sensors <- bind_rows(obs$sensors, d |> filter(valid_day, !is.na(timepoint)) |>
                                 distinct(study_id, timepoint))
    }
  }
}
for (nm in names(sources)) id_audit_line(sources[[nm]], nm)
```

## ID audit across modalities

```{r}
#| label: id-audit
log_section("ID audit")
sets <- build_id_source_sets(sources)
fmt <- id_format_inventory(sets)
readr::write_csv(fmt, file.path(validation_dir, paste0(stem, "_modality_id_formats.csv")))
if (dplyr::n_distinct(fmt$len_min) > 1 || dplyr::n_distinct(fmt$len_max) > 1) {
  cat("⚠️ ID length differs between sources; a join will under-match. Fix id_transform in the config.\n")
}
fmt
overlap <- id_overlap_summary(sets)
only <- id_only_in_each(sets)
readr::write_csv(only, file.path(validation_dir, paste0(stem, "_modality_id_only_in_each.csv")))
cat("• ", nrow(only), " ID(s) appear in exactly one source; adjudicate each\n", sep = "")
overlap
```

## The participant × timepoint index

```{r}
#| label: index
#| fig-width: 10
#| fig-height: 6
log_section("Index")
index <- tidyr::expand_grid(study_id = sort(unique(roster$study_id)), timepoint = sched$timepoint)
cov_list <- list()
for (m in intersect(modalities, names(obs))) {
  o <- obs[[m]] |> mutate(study_id = as.character(study_id), timepoint = as.character(timepoint), .obs = TRUE)
  index <- index |> left_join(o |> distinct(study_id, timepoint, .obs), by = c("study_id", "timepoint")) |>
    mutate(!!paste0("has_", m) := coalesce(.obs, FALSE)) |> select(-.obs)
  cov_list[[m]] <- timepoint_coverage(obs[[m]], "study_id", "timepoint", config, modality = m,
                                      expected_ids = roster$study_id)
}
index <- index |> mutate(timepoint = factor(timepoint, levels = sched$timepoint)) |> arrange(study_id, timepoint)
assert_expected_n(index, dplyr::n_distinct(roster$study_id), id_col = "study_id", label = "index build")
write_versioned_rds(index, paste0(stem, "_multimodal_index"), derived_dir)
write_latest_rds(index, paste0(stem, "_multimodal_index"), derived_dir)
cat("✓ Wrote ", stem, "_multimodal_index_latest.rds (", nrow(index), " rows)\n", sep = "")

dash <- modality_coverage_dashboard(lapply(cov_list, timepoint_coverage_summary))
readr::write_csv(dash, file.path(validation_dir, paste0(stem, "_modality_coverage.csv")))
if (requireNamespace("ggplot2", quietly = TRUE)) {
  p <- plot_coverage_dashboard(dash)
  ggplot2::ggsave(file.path(figures_dir, paste0(stem, "_modality_coverage.png")), p, width = 10, height = 3.5, dpi = 120)
  print(p)
  print(plot_timepoint_completeness(bind_rows(cov_list)))
}
dash
```

## Linked analysis table

```{r}
#| label: link
log_section("Link")
linked <- index
if (!is.null(eeg)) {
  linked <- linked |> left_join(eeg |> mutate(timepoint = factor(timepoint, levels = sched$timepoint)),
                                by = c("study_id", "timepoint"))
}
if ("sensors" %in% modalities) {
  for (sk in names(config$sensors$streams)) {
    d <- latest(paste0("sensor_", sk, "_day_valid"))
    if (is.null(d) || !"timepoint" %in% names(d)) next
    metrics <- names(config$sensors$streams[[sk]]$metrics)
    agg <- d |> filter(valid_day, !is.na(timepoint)) |>
      summarise(across(all_of(metrics), ~ mean(.x, na.rm = TRUE)), n_valid_days = n(),
                .by = c(study_id, timepoint)) |>
      rename_with(~ paste0(sk, "_", .x), -c(study_id, timepoint)) |>
      mutate(timepoint = factor(timepoint, levels = sched$timepoint))
    linked <- linked |> left_join(agg, by = c("study_id", "timepoint"))
  }
}
assert_expected_n(linked, nrow(index), id_col = NULL, label = "modality joins (rows)")
write_versioned_rds(linked, paste0(stem, "_multimodal_linked"), derived_dir)
write_latest_rds(linked, paste0(stem, "_multimodal_linked"), derived_dir)
cat("✓ Wrote ", stem, "_multimodal_linked_latest.rds: ", ncol(linked), " columns\n", sep = "")
```

## What happens next

07c runs the full UpSet audit on the same sources; 08 and 10 read
`<stem>_multimodal_linked_latest.rds`. Approve this stage when the ID format
table shows one length and one shape across sources, the only-in-each list is
one you can account for participant by participant, and the coverage
dashboard shows the pattern you expect from the study's calendar.

```{r}
#| label: record
builder_record_render("07m_link_modalities")
```
